### 2016-07-07 image_setup.tex
Image Setup Section done!

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Notice that \texttt{set\_hostname} is now called by the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at next system startup.

\subsection{Image Package Updating}
Besides adding and configuring files within the image, the testbed
administrator can also install and update the software packages within
the image which will be illustrated in this subsection.

It may be desired to pre-install some programs in the image before it is
written to all the memory cards that goes into the \ac{Raspi}s.
This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
(change root).
%\url{https://wiki.debian.org/QemuUserEmulation}
%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}

Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
install some additional software:
%Because \ac{Raspi}s are equipped with an \ac{ARM} processor
%%it is not as
%straightforward as performing the task for two Linux systems of the same
%architecture.

@@

Make sure the second command writes "enabled" as the output above. If that
is not the case, then try enabling it:

@@

Provided that qemu-arm is enabled, we should now be able to change root (chroot)
into our Raspbian lite image. Change root is a method in Linux that enables
the root (/) to be changed. Thus, enabling a Linux installation within a
Linux installation.

There are a few commands to be performed before actually changing root into the
root partion of the image.
First, to get internet access from within the Raspbian lite image it is needed
to copy our \texttt{resolv.conf} into the filesystem. Line 1 change directory
into root filesystem of the image. The second line copies \texttt{resolv.conf}
from the testbed administrators \ac{PC} to the image.

% CHROOT to OS image
\begin{lstlisting}[]
$ cd $ROOTDIR
# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, because of the \ac{ARM} architecture,
\texttt{/usr/bin/qemu-arm-static} needs to be copied into the image before continuing:

@@

The final preparation before changing root is to populate the directories \texttt{proc},
\texttt{sys}, and \texttt{dev}:

@@

Finally, it is time to change root.
%Finally, the filesystem is ready for us to change root.
This can be done with \texttt{proot}. It may be required to install
\texttt{proot} using \texttt{apt-get install proot}:

% CHROOT to OS image
\begin{lstlisting}[]
# proot -q qemu-arm-static -S ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Most available material online uses the more known \texttt{chroot} command as
written in the code block below instead of \texttt{proot}. This did not work
correctly in our machines, but we present the command as an alternative to
\texttt{proot} in case the testbed administrator should be in the opposite
situration where \texttt{proot} is not working.

% CHROOT to OS image
\begin{lstlisting}[]
# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If everything went well, we should now be \texttt{chrooted} into the Raspbian lite
filesystem under its root. Optionally, change the prompt title to indicate that it
is a \texttt{chrooted} environment:

@@

The Raspbian lite system image should now be possible to use almost as if it
had been booted in a \ac{Raspi}.
A major difference is that the testbed administrator's \ac{PC} is likely
significantly faster than a \ac{Raspi}.
%The only difference is that the laptop that we are using is
%significantly faster than a \ac{Raspi}.
Thus enabling updates, upgrades, and installing new software packages much
faster than in the \ac{Raspi}s.
%Thus enabling us to for example update,
%upgrade, and install software packages:

@@

We also install a few packages that we find useful:
%Lets install some useful applications:
% Install packages
\begin{lstlisting}[]
(chroot) # apt-get install vim git screen
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

When writing the image to a memory card, all these change that has been made
to the image so far will exists in all \ac{Raspi}s.
%All the changes that are made here will exists in all \ac{Raspi}s when the
%image is written to a memory card.

>>>>

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Notice that \texttt{set\_hostname} is now called by the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at next system startup.

%\url{https://wiki.debian.org/QemuUserEmulation}
%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
\subsection{Image Package Updating}
Besides adding and configuring files within the image, the testbed
administrator may want to install and update the software packages
within the image before it is written to all the memory cards that goes
into the \ac{Raspi}s. From any Linux x86 machine as the testbed administrator
\ac{PC}, this can be done using \texttt{chroot} command in the
QEMU~\cite{QemuUserEmulation} hypervisor for \ac{ARM} processors.

\texttt{chroot} is a method in Linux that modifies the apparent root
filesystem location from \texttt{/} to any other path. Consequently, in
our case, we can use the Raspbian Lite image root filesystem within the
testbed administrator Linux distribution. Then, QEMU allows the execution
of commands for the \ac{Raspi} image (\ac{ARM} instructions) through the ones
from the testbed administrator \ac{PC} architecture. Due to the \ac{ARM}
processor that the \ac{Raspi}s employ, t is required to install the QEMU related software first and verigy that QEMU is \ac{ARM} enabled. To do so,
run the following commands:

@@

In the previous output, the testbed administrator must be sure that the second
command writes \texttt{qemu-arm (enabled)} as indicated. If that is not the
case, then it should be possible to enable it by running:

@@

Provided that qemu-arm is enabled, we should now be able to \texttt{chroot}
into our Raspbian lite image. However, there are a few commands to be
performed before actually changing root into the root partion of the image.
First, to get internet access from within the Raspbian lite image it is needed
to copy the testbed administrator Linux distribution \texttt{resolv.conf} file
into the image root filesystem. To do this, it is required to run the
following:

\begin{lstlisting}[]
$ cd $ROOTDIR
# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, because of the \ac{ARM} architecture, the
\texttt{/usr/bin/qemu-arm-static} command needs to be copied into the
image before continuing by running:

@@

Before changing root it is necessary to populate the directories
\texttt{proc}, \texttt{sys} and \texttt{dev} for the image to get control
as the testbed administrator apparent root filesystem. This is made by the
following commands:

@@

Finally, to change root the following needs to be run:

\begin{lstlisting}[]
# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

% In some cases, \texttt{chroot} may not work properly. tgus, in this case, an
% alternative can be \texttt{proot}. It may be required to install it
% before changing the root with \texttt{apt-get install proot}. Then, it
% should possible for the testbed administrator to run the following:

% % CHROOT to OS image
% \begin{lstlisting}[]
% # proot -q qemu-arm-static -S ${ROOTDIR}
% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}
If successfully executed, our terminal should have changed the prompt
indicating that we are the root user in the Raspbian lite root filesystem as
the apparent root. To be aware of the mode that we are working now, we change
the prompt title to indicate that it is a \texttt{chroot} environment as
follows:

@@

The Raspbian lite image should now be possible to use almost as if it had
been booted in a \ac{Raspi}. A major difference is that the testbed
administrator \ac{PC} is likely significantly faster than a \ac{Raspi}.
Hence, enabling updates, upgrades and installing new software packages
should be much faster than in a \ac{Raspi}. Still updating and upgrading
the packages for the \ac{Raspi} might take some amount of time. To get
the upgraded system packages, it is necessary to run:

@@

We further install some packages that we consider useful:
%Lets install some useful applications:
% Install packages
\begin{lstlisting}[]
(chroot) # apt-get install vim git screen
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
for managing Git repositories and \texttt{screen} for better handling of
long-runnning processes. When writing the image either to a memory card
or storing it remotely \todot{Are we going to store the image
into a memory card or remotely in a URL? Since this is not completely clear
yet.}, all these change that has been made to the image so far will exists in
all \ac{Raspi}s.
